Experience Report: Haskell in Computational Biology

### 2012-06-03 mrfy_experience_report.tex
sketch of what is wrong with Norman's code

@@

MRFy's search computes a sequence of \emph{generations}.
A~generation is a collection of scored {placements}.
We~pass the
search function a \texttt{Scorer}, which runs Viterbi's algorithm.
\smallverbatiminput{scoredecl}
A~search heuristic is characterized by four functions, which are
bundled into a record called \texttt{SearchStrategy}.
\smallfuzzverbatiminput{2.3pt}{strategy}
In this record,
\begin{itemize}
\item
Function \texttt{gen0} produces the first generation of placements.
\item
Function \texttt{nextGen} produces a new generation of placements from
an existing generation.
Because scoring a placement is expensive, 
and some placements may persist from the old generation to the new
generation, \texttt{nextGen} works with \emph{scored} placements, not
bare placements.
\item
A~new generation produced by \texttt{nextGen} may be no better
than the generation it replaces.
When it is first produced, the new generation is ``on trial.''
The \texttt{accept} function decides whether to keep a new generation
or to try \texttt{nextGen} again with a different seed.
The \texttt{accept} function considers a list of type
\texttt{[Scored~Age]}.
At~the head of this list is the score of the best
placement of the trial generation, together with the \texttt{Age} of the trial
generation, which is the total number of generations tried.
The remaining elements of the list hold the history of scores and ages for
previously \emph{accepted} generations.
(The list omits generations that didn't survive the \texttt{accept} trial.)
For convenience, the \texttt{accept} function also receives
the \texttt{Age} of the trial generation.
\item
Finally, the \texttt{quit} function decides when to stop the search.
\ifnotcutting
It~too considers the list of accepted generations and the age of the current
generation. 
\fi
\end{itemize}
\seclabel{search}%
We hope you find the search simple and straightforward:
\smallfuzzverbatiminput{11pt}{search}

>>>>

\newcommand\mrfy{MRFy} % damn you Noah

@@

MRFy's search computes a sequence of \emph{generations}.
A~generation is a collection of scored {placements}.
\begingroup\em\relax
Ideally, MRFy's search would use now-classic lazy, modular
technique advocated by \citet{hughes:why}, in which one function
computes an infinite sequence of generations, and another function
uses a finite prefix to decide on an approximation.
But making MRFy's search lazy and modular search is not so easy.
 because MRFy's search is stochastic.

In~a stochastic search, \emph{movement} between points in the search
space is not the same as \emph{progress}.
For example, suppose you are searching for the top of a hill
and you take a small jump in a random direction.
If~you arrive at a higher point than your point of origin, you have
made progress.
Otherwise you have just moved.
\smallverbatiminput{movequality}
We~hoped to implement \mrfy's search by composing functions:
\begin{itemize}
\item
From a given point, a \emph{generator} produces an infinite sequence
of moves.
\item
From such a sequence, a \emph{classifier} chooses the first move that
represents progress.
What sort of move constitutes progress is not only different for
different kinds of searches, but may also be different at different
\emph{times} in a search.
\end{itemize}
A~generator and classifier can be composed to produce a sequence of
progressing generations, and a test function can terminate the search
after examining a finite
prefix.
But this sketch contains a fatal flaw:
a generator may produce an infinite sequence of generations without any
progress.
And then, when the test function tries to examine a finite prefix of
progressing generations, it will diverge.
We~might conclude that
\mrfy's search algorithm is inherently non-modular:
production of progressing generations cannot be decoupled from the
test for termination.

N.B.~The following transition no longer makes any sense:
\endgroup
We~pass the
search function a \texttt{Scorer}, which runs Viterbi's algorithm.
\smallverbatiminput{scoredecl}
A~search heuristic is characterized by four functions, which are
bundled into a record called \texttt{SearchStrategy}.
\smallfuzzverbatiminput{2.3pt}{strategy}
In this record,
\begin{itemize}
\item
Function \texttt{gen0} produces the first generation of placements.
\item
Function \texttt{nextGen} produces a new generation of placements from
an existing generation.
Because scoring a placement is expensive, 
and some placements may persist from the old generation to the new
generation, \texttt{nextGen} works with \emph{scored} placements, not
bare placements.
\item
A~new generation produced by \texttt{nextGen} may be no better
than the generation it replaces.
When it is first produced, the new generation is ``on trial.''
The \texttt{accept} function decides whether to keep a new generation
or to try \texttt{nextGen} again with a different seed.
The \texttt{accept} function considers a list of type
\texttt{[Scored~Age]}.
At~the head of this list is the score of the best
placement of the trial generation, together with the \texttt{Age} of the trial
generation, which is the total number of generations tried.
The remaining elements of the list hold the history of scores and ages for
previously \emph{accepted} generations.
(The list omits generations that didn't survive the \texttt{accept} trial.)
For convenience, the \texttt{accept} function also receives
the \texttt{Age} of the trial generation.
\item
Finally, the \texttt{quit} function decides when to stop the search.
\ifnotcutting
It~too considers the list of accepted generations and the age of the current
generation. 
\fi
\end{itemize}
\seclabel{search}%
We hope you find the search simple and straightforward:
\smallfuzzverbatiminput{11pt}{search}

\remark{We need to explain why we don't just generate an infinite lazy
list of thingies.  And perhaps we can consider an infinite tree!}
